---
title: "NYC K-5 Chronic Absenteeism Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}

# load required packages
library(flexdashboard)
library(data.table)
library(tidyverse)
library(broom)
library(lubridate)
library(plotly)
library(shiny)

# load saved data
df_absent <- readRDS("df_absent.rds")
```

# Overview

**Add text here about the project.** 

Code for help with creating this dashboard here: https://epirhandbook.com/new_pages/flexdashboard.html

**Working On**

total number of students in category
rates total, by borough, by school district (new variable)

by grade would also be cool 


```{r}

# a few modification to the data frame 
df_absent <- df_absent %>% 
  mutate(Year = ymd(paste0(substr(Year, 1, 4), "-01-01")), # change to data format, taking first year (lubridate)
         chronic_absent_pct = round(chronic_absent_pct,2)) # round chronic absenteeism rate

# create graph template to use 

template_plot <- function(df, demographic_category) {
  plot <- df %>%
    filter(Demographic.Category == demographic_category) %>% # specify which demographic cat below in renderPlotly()
    ggplot(aes(text = paste(
      "<b> Year: ", Year, "<br>",
      "Percent Chronically Absent: ", chronic_absent_pct, "<br>", # this is what we want displayed in the dynamic plot
      "Variable: ", Demographic.Variable))) +
    geom_line(aes(x = Year, y = chronic_absent_pct, color = Demographic.Variable), inherit.aes = FALSE) +
    geom_point(aes(x = Year, y = chronic_absent_pct, color = Demographic.Variable)) +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    scale_y_continuous(
      limits = c(0, 60),
      expand = c(0, 0)
    ) +
    labs(
      title = "",
      x = "Year",
      y = "Chronic Absenteeism (%)",
      color = ""
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom",
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.title = element_text(size = 10),
      legend.text = element_text(size = 9)
    )
  
  ggplotly(plot, tooltip = "text") # tooltip uses the text above for hover-over info
}


```


# All Students{data-navmenu="Demographic Indicator"}

### Chronic Absenteeism Rates Over Time for All Students

```{r}

# use this function with different "Demographic.Variable" to render new dynamic plot from the template created above

renderPlotly({
  template_plot(df_absent, "All Students")
})

```


# Ethnicity{data-navmenu="Demographic Indicator"}

### Chronic Absenteeism Rates Over Time by Ethnicity

```{r}

renderPlotly({
  template_plot(df_absent, "Ethnicity")
})

```


# English Language Learners Status{data-navmenu="Demographic Indicator"}

### Chronic Absenteeism Rates Over Time by ELL Status

```{r}

renderPlotly({
  template_plot(df_absent, "ELL Status")
})

```

# Gender{data-navmenu="Demographic Indicator"}

### Chronic Absenteeism Rates Over Time by Gender

```{r}

renderPlotly({
  template_plot(df_absent, "Gender")
})

```

# Poverty{data-navmenu="Demographic Indicator"}

### Chronic Absenteeism Rates Over Time by Poverty Status

```{r}

renderPlotly({
  template_plot(df_absent, "Poverty")
})

```

# Students with Disability{data-navmenu="Demographic Indicator"}

### Chronic Absenteeism Rates Over Time by Students with Disability Status

```{r}

renderPlotly({
  template_plot(df_absent, "SWD Status")
})

```

# Students in Temporary Housing{data-navmenu="Demographic Indicator"}

### Chronic Absenteeism Rates Over Time by Students in Temporary Housing Status

```{r}

renderPlotly({
  template_plot(df_absent, "STH Status")
})

```

# Borough{data-navmenu="Demographic Indicator"}

# Grade{data-navmenu="Demographic Indicator"}